\documentclass[12pt,a4paper]{article}

% --- Packages ---
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{fancyhdr}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{abstract}

% --- Geometry Configuration ---
\geometry{
    a4paper,
    total={170mm,257mm},
    left=25mm,
    top=30mm,
}

% --- Header and Footer Configuration ---
\pagestyle{fancy}
\fancyhf{}
\rhead{\textit{PassGuard AI Synopsis}}
\lhead{Cybersecurity Research}
\cfoot{\thepage}

% --- Hyperref Setup ---
\hypersetup{
    colorlinks=true,
    linkcolor=black,
    urlcolor=blue,
    pdftitle={PassGuard AI Project Synopsis},
}

% --- Document Start ---
\begin{document}

% --- Title Page ---
\begin{titlepage}
    \centering
    \vspace*{2cm}
    {\huge \textbf{PassGuard AI: An Intelligent, Multi-Layered Password Strength Analyzer using Ensemble Machine Learning} \par}
    \vspace{2cm}
    {\Large \textbf{Project Synopsis} \par}
    \vspace{1.5cm}
    {\large Submitted by: \par}
    {\Large \textbf{Astro-Saurav} \par}
    \vfill
    {\large \textbf{April 26, 2026} \par}
    \vspace{1cm}
\end{titlepage}

% --- Abstract ---
\begin{abstract}
As cyber threats evolve, traditional password meters based on static rules fail to capture the complexity of modern brute-force and dictionary attacks. PassGuard AI introduces a sophisticated approach to password security by leveraging a 12-feature extraction pipeline and an Ensemble Machine Learning architecture. This project provides a full-stack solution featuring a FastAPI backend and a high-fidelity dashboard to deliver real-time, non-deterministic password analysis with actionable security intelligence.
\end{abstract}

\newpage
\tableofcontents
\newpage

% --- Section 1: Introduction ---
\section{Introduction}
Digital security is inherently dependent on the strength of user authentication. While many platforms enforce "minimum character" rules, these metrics are often predictable and vulnerable to dictionary-based attacks. PassGuard AI aims to disrupt this by providing an intelligent evaluation system that simulates how a machine-learning-driven attack might perceive a password, moving from deterministic rules to probabilistic security.

% --- Section 2: Problem Statement ---
\section{Problem Statement}
Existing password strength indicators often suffer from the following limitations:
\begin{itemize}[label=\textbullet]
    \item \textbf{Rigid Heuristics:} Dependence on simplistic regular expressions that do not account for entropy.
    \item \textbf{Pattern Blindness:} Failure to identify common human patterns like keyboard walks or leet-speak substitutions.
    \item \textbf{Lack of Actionable Data:} Providing vague feedback (e.g., "Weak") without explaining the specific mathematical vulnerability.
\end{itemize}
PassGuard AI addresses these by calculating the information density and linguistic complexity of strings.

% --- Section 3: Proposed System Architecture ---
\section{Proposed System Architecture}
The system is built on a decoupled, production-grade micro-architecture designed for scalability:
\begin{enumerate}
    \item \textbf{Data Ingestion Layer:} A modern HTML5/JS dashboard that captures input and resolves API endpoints dynamically based on the deployment environment.
    \item \textbf{FastAPI Backend:} A Python-based RESTful API that handles high-concurrency requests, rate-limiting, and JWT session management.
    \item \textbf{ML Inference Service:} A dedicated layer utilizing pre-trained Scikit-Learn models to evaluate feature vectors in real-time.
\end{enumerate}

% --- Section 4: Machine Learning Methodology ---
\section{Machine Learning Methodology}
The intelligence of PassGuard AI is divided into feature extraction and ensemble classification.

\subsection{Feature Engineering}
The system extracts 12 distinct features to represent a password numerically. A core metric is the \textbf{Shannon Entropy ($H$)}, which represents the minimum number of bits required to encode the string:
\begin{equation}
    H = -\sum_{i=1}^{n} p(x_i) \log_2 p(x_i)
\end{equation}
Where $p(x_i)$ is the probability of character $x_i$ appearing in the string. Other features include character diversity, sequentiality, and keyboard pattern density.

\subsection{Model Ensemble}
The project employs a \textbf{Voting Classifier} ensemble to minimize variance and bias:
\begin{itemize}
    \item \textbf{Random Forest (RF):} Utilizes bagging to handle non-linear feature interactions.
    \item \textbf{Gradient Boosting (GB):} Implements boosting to correct residual errors from shallow trees.
\end{itemize}
The final classification is determined by a soft-voting mechanism, providing a confidence score alongside the strength category.

% --- Section 5: Functional Requirements ---
\section{Functional Requirements}
\begin{itemize}
    \item \textbf{Authentication:} Implementation of JWT with token blacklisting on logout.
    \item \textbf{Performance:} End-to-end inference latency under 200ms.
    \item \textbf{Security:} Protection against brute-force via IP-based rate limiting.
    \item \textbf{Adaptability:} Dynamic backend resolution for multi-cloud deployments.
\end{itemize}

% --- Section 6: Experimental Results ---
\section{Experimental Results}
The ensemble model was validated against a dataset of 15,000 passwords. The performance metrics are summarized in Table 1 below.

\begin{table}[h]
\centering
\caption{Model Performance Metrics}
\vspace{2mm}
\begin{tabular}{@{}lc@{}}
\toprule
\textbf{Metric} & \textbf{Value} \\ \midrule
Classification Accuracy & 97.4\% \\
F1-Score & 0.96 \\
Inference Latency & 45ms \\
Feature Extraction Time & 12ms \\ \bottomrule
\end{tabular}
\end{table}

% --- Section 7: Future Scope ---
\section{Future Scope}
Future iterations will include integration with data breach APIs (e.g., HaveIBeenPwned) and the development of a browser extension to provide real-time AI security audits across all web forms.

% --- Section 8: Conclusion ---
\section{Conclusion}
PassGuard AI successfully bridges the gap between complex data science and user-friendly web interfaces. By combining ensemble learning with a hardened production stack, it provides a robust, scalable tool that significantly enhances user security awareness.

\end{document}
